The Chemistry Development Kit (CDK). 3. Atom typing, Rendering, Molecular Formula, and Substructure Searching

### 2014-06-25 article.tex
Cleaned up the background/introduction

Open Source cheminformatics has made significants steps recently~\cite{OBoyle2011}.
Development of the Chemistry Development Kit (CDK) has been previously documented~\cite{Steinbeck2003,Steinbeck2006},
but followed this general trend in open source cheminformatics, and increased in functionality
and adoption. For example, new tools expose CDK functionality, such as Cinfony~\cite{OBoyle2008},
rcdk~\cite{}, and workflow plugins for Taverna\cite{Truszkowski2011}, KNIME\cite{Beisken2013}.
Furthermore, projects with specific functionality have emerged building on the CDK, including
jCompoundMapper~\cite{Hinselmann2011}, ScaffoldHunter~\cite{wetzel2009interactive}, OMG~\cite{Peironcely2012},
Padel~\cite{yap2011padel}, AMBIT-SMARTS~\cite{jeliazkova2011ambitsmarts}, AMBIT-Tautomer~\cite{kochev2013ambit},
SMSD~\cite{}, and FooFromRydberg~\cite{}. Other tools used previous versions already, and follow
newer releases of the CDK, such as Bioclipse~\cite{} and AMBIT~\cite{jeliazkova2011ambit}.

Importantly, the CDK has seen independent validation ... of both functionality, quality, and performance...
... fingerprint results ... performance paper ... Sayle MDL atom type work.

>>>>

Open Source cheminformatics has made significants steps forward recently~\cite{OBoyle2011}.
The Chemistry Development Kit (CDK) is one of the tools under the Blue Obelisk wings,
and previously documented version have been widely adopted~\cite{Steinbeck2003,Steinbeck2006}.
For example, new tools expose CDK functionality, such as Cinfony~\cite{OBoyle2008},
rcdk~\cite{}, and workflow plugins for Taverna\cite{Truszkowski2011} and KNIME\cite{Beisken2013}.
Furthermore, projects with specific functionality have emerged building on the CDK, including
jCompoundMapper~\cite{Hinselmann2011}, ScaffoldHunter~\cite{wetzel2009interactive}, OMG~\cite{Peironcely2012},
Padel~\cite{yap2011padel}, AMBIT-SMARTS~\cite{jeliazkova2011ambitsmarts}, AMBIT-Tautomer~\cite{kochev2013ambit},
and SMSD~\cite{}. Other tools used previous versions already, and follow
newer releases of the CDK, such as Bioclipse~\cite{} and AMBIT~\cite{jeliazkova2011ambit}.

However, previous version of the CDK were written by a community with specific applications
in mind, of which structure elucidation was one. It could therefore be rather competitive
in some cheminformatics domains, such as fingerprinting~\cite{Ekins,Cannon}, in other parts
it could be inadequate, such in the capturing of stereochemistry.

Continued community development of the CDK in the last few years, however, has shown that
Open Source software is here to stay. The adoption of automatic build systems and
quality control methodologies such as unit testing, automated source code validation,
and peer review by fellow developers have greatly improved the stability of the library,
At the expense of slowing down the development, it has allowed for cleaning up interdependencies
between modules of code, and has reset the scalability of the development model,
allowing for a lot of new functionality in code APIs while maintaining the quality
of code depending on those core APIs.

This has resulted in important improved functionality, such as a separately published
InChI functionality~\cite{SpjuthXXXX} and greatly improved ring detection functionality~\cite{MayYYYY},
but also a core atom type perception module, covering a much wider set of elements than
previous CDK versions and capturing many more charged and radical atom species,
a more comprehensive fingerprinting API, many speed and stability improvements, and more.
